FITS format for planetary surfaces: definitions, applications and best practices.

### 2018-02-07 ess_marmo.tex
Update on Overleaf.

 \authors{C. Marmo\affil{1},
 T. M. Hare\affil{2},
 S. Erard\affil{3},
 B. Cecconi\affil{3},
 M. Minin\affil{4},
 A. Zinzi\affil{5},
 A. P. Rossi\affil{4}}

\affiliation{1}{GEOPS, Univ. Paris-Sud, CNRS, Univ. Paris-Saclay,
Orsay, France}
\affiliation{2}{U. S. Geological Survey, Astrogeology Science Center,
Flagstaff, AZ, USA}
\affiliation{3}{LESIA, Observatoire de Paris, PSL Research University,
CNRS, Sorbonne Universit\'es, UPMC Univ. Paris 06, Univ. Paris Diderot,
Sorbonne Paris Cit\'e, Meudon, France}
\affiliation{4}{Jacobs University, Bremen, Germany}
%\affiliation{5}{}

@@

Assuming projected products distributed in FITS format a quick method to add GIS support for
FITS images is to use a GDAL-supported detached header.
GDAL Virtual Header files simply describe the internal structure of a FITS image, but
also the map projection and body size in a standardized "well known text"
(WKT) projection string.
As a result, the FITS raster loaded into QGIS thanks to its correspondent GDAL Virtual Header
is indistinguishable from the GeoTIFF layer (see figure \ref{fig:ds9qgis}, right side).

MATISSE\footnote{http://tools.asdc.asi.it/matisse.jsp} \citep{ZINZI2016}

@@

\subsubsection{OMEGA spectra and geometry data}

\subsubsection{Akatsuki imaging data}
\section{Inter-operable software developments}
\label{sec:softdev}

@@

\acknowledgments
This work benefits from support of VESPA/Europlanet.
The Europlanet 2020 Research Infrastructure is funded by the European Union
under the Horizon 2020 research and innovation program, grant agreement N.654208.

>>>>

 \authors{C. Marmo\affil{1},
 T. M. Hare\affil{2},
 S. Erard\affil{3},
 M. Minin\affil{4},
 A. Zinzi\affil{5,6},
 B. Cecconi\affil{3},
 A. P. Rossi\affil{4}}

\affiliation{1}{GEOPS, Univ. Paris-Sud, CNRS, Univ. Paris-Saclay,
Orsay, France}
\affiliation{2}{U. S. Geological Survey, Astrogeology Science Center,
Flagstaff, AZ, USA}
\affiliation{3}{LESIA, Observatoire de Paris, PSL Research University,
CNRS, Sorbonne Universit\'es, UPMC Univ. Paris 06, Univ. Paris Diderot,
Sorbonne Paris Cit\'e, Meudon, France}
\affiliation{4}{Jacobs University, Bremen, Germany}
\affiliation{5}{ASI Science Data Center, c/o ASI, Via del Politecnico snc, 00133, Rome, Italy}
\affiliation{6}{INAF-OAR, Via Frascati n. 33, 00078, Monte Porzio Catone (RM), Italy}

@@

Assuming projected products distributed in FITS format, a quick method to add GIS support for
FITS images is to use a GDAL-supported detached header.
GDAL Virtual Header files simply describe the internal structure of a FITS image, but
also the map projection and body size in a standardized "well known text"
(WKT) projection string.
As a result, the FITS raster loaded into QGIS thanks to its correspondent GDAL Virtual Header
is indistinguishable from the GeoTIFF layer (see figure \ref{fig:ds9qgis}, right side).

In order to straightforward load FITS files in GIS software, it is recommended to have
body radii and linear coordinates filled in.
Information about invalid values (NoData values) and minimum and maximum values are useful
for correct display.

\subsubsection{OMEGA spectra and geometry data}
As stressed above, having planetary data distributed in FITS at any processing level is
a way to avoid multiple conversions in the processing chain and simplify reprocessing.
As an example of non projected planetary data we have converted in FITS format data
acquired by the spectroimager OMEGA onboard of Mars Express.
OMEGA unprojected and uncalibrated data are distributed by ESA-PSA and PDS archives.
Their structure is described in the OMEGA Experiment Archive 
Interface Control Document\footnote{ftp://psa.esac.esa.int/pub/mirror/MARS-EXPRESS/OMEGA/MEX-M-OMEGA-2-EDR-FLIGHT-V1.0/DOCUMENT/EAICD\_OMEGA.PDF}.
In planetary science a first spatial calibration data often consists in providing coordinates
on specific points on the detector (the center of the pixel, one or more pixel corners, etc.)

\subsubsection{Akatsuki imaging data}

@@

\section{Inter-operable software developments}
VO-QGIS-plugin\footnote{https://github.com/epn-vespa/VO\_QGIS\_plugin}
MATISSE\footnote{http://tools.asdc.asi.it/matisse.jsp} \citep{ZINZI2016}
\label{sec:softdev}

@@

\acknowledgments
This work benefits from support of VESPA/Europlanet.
The Europlanet 2020 Research Infrastructure is funded by the European Union
under the Horizon 2020 research and innovation program, grant agreement N.654208.
C. Marmo would lik to thanks Mark Calabretta from Australia Telescope National Facility
for his WCSLIB library, and David Berry from Joint Astronomy Center, Hilo, for useful
discussions about TAB projection in FITS.
